A Simplified Spectral Deferred Correction Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium

### 2020-08-16 simplified_sdc_intro/paper.tex
update NSE flow

For this study we will use the 19 isotope ``aprox19'' network together
with a table that gives the NSE abundances in regions where the system
is in NSE.  The table provides:
\begin{align*}
Y_e &= \sum_k \frac{Z_k X_k}{A_k} \\
\bar{A} &= \left [ \sum_k \frac{X_k}{A_k} \right ]^{-1} \\
\frac{B}{A} &= \sum_k \frac{B_k X_k}{A_k}
\end{align*}
where $B_k$ is the binding energy of nucleus $k$.  In our simulations,
we store these 3 quantities as auxiliary data that is carried along
with the rest of the fluid state in each zone.

For Strang split coupling of hydro and reactions, $DX_k/Dt = 0$,
and our evolution equations are:
\begin{align*}
\frac{DY_e}{Dt} &= \sum_k \frac{Z_k}{A_k} \frac{DX_k}{Dt} = 0 \\
\frac{D}{Dt} \frac{1}{\bar{A}} &= - \frac{1}{\bar{A}^2} \frac{D\bar{A}}{Dt} = \sum_k \frac{1}{A_k} \frac{DX_k}{Dt} = 0 \rightarrow \frac{D\bar{A}}{Dt} = 0 \\
\frac{D}{Dt} \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k}{A_k} \frac{DX_k}{Dt} = 0
\end{align*}
Therefore each of these auxillar equations obeys an advection
equation in the hydro part of the advancement.  In the SDC algorithm,
there will be a reactive source (an $\Ic_q$) for each of these that is
computed in the same manner as above.

@@

For Strang splitting, the basic flow of using the aprox19 + NSE network
is as follows:
\begin{itemize}
\item initialize the problem, including $X_k$

\item fill the initial aux data with $Y_e$, $\bar{A}$, and $(B/A)$.

\item in hydro, update these quantities simply via advection

\item for the reactive update:

  \begin{itemize}
  \item check if NSE applies (see below)

  \item if we are in an NSE region:

    \begin{itemize}
    \item use $\rho$, $T$, and $Y_e$ to call the table. This
      returns: $dY_e/dt$, $(B/A)_{\rm out}$, and $\bar{A}_{\rm out}$.

    \item update $Y_e$:
    \begin{equation}
      (Y_e)_{\rm out} = (Y_e)_{\rm in} - \Delta t \frac{dY_e}{dt}
    \end{equation}

  \item update the auxiliary state value of $\bar{A}$ to be simply
    $\bar{A}_{\rm out}$ from the table.

@@

  \item update the mass fractions, $X_k$, using the values from the table
    \end{itemize}

  \item if we are not in NSE:

@@

We use the following conditions to determine if we are in NSE:
\begin{itemize}
\item $\rho > \rho_\mathrm{nse}$
\item $T > T_\mathrm{nse}$
\item $X(\isotm{C}{12}) < A_\mathrm{nse}$
\item $X(\isotm{He}{4}) + X(\isotm{Cr}{48}) + X(\isotm{Fe}{52}) + X(\isotm{Fe}{54}) + X(\isotm{Ni}{56}) > B_\mathrm{nse}$
\end{itemize}

>>>>

For this study we will use the 19 isotope ``aprox19'' network together
with a table that gives the NSE abundances in regions where the system
is in NSE.  The table provides:
\begin{align}
Y_e &= \sum_k \frac{Z_k X_k}{A_k} \\
\bar{A} &= \left [ \sum_k \frac{X_k}{A_k} \right ]^{-1} \\
\frac{B}{A} &= \sum_k \frac{B_k X_k}{A_k}
\end{align}
where $B_k$ is the binding energy of nucleus $k$.  In our simulations,
we store these 3 quantities as auxiliary data that is carried along
with the rest of the fluid state in each zone.

For Strang split coupling of hydro and reactions, $DX_k/Dt = 0$,
and our evolution equations are:
\begin{align}
\frac{DY_e}{Dt} &= \sum_k \frac{Z_k}{A_k} \frac{DX_k}{Dt} = 0 \\
\frac{D}{Dt} \frac{1}{\bar{A}} &= - \frac{1}{\bar{A}^2} \frac{D\bar{A}}{Dt} = \sum_k \frac{1}{A_k} \frac{DX_k}{Dt} = 0 \rightarrow \frac{D\bar{A}}{Dt} = 0 \\
\frac{D}{Dt} \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k}{A_k} \frac{DX_k}{Dt} = 0
\end{align}
Therefore each of these auxillar equations obeys an advection
equation in the hydro part of the advancement.  In the SDC algorithm,
there will be a reactive source (an $\Ic_q$) for each of these that is
computed in the same manner as above.

@@

For Strang splitting, the basic flow of using the aprox19 + NSE network
is as follows:
\begin{itemize}
\item {\em Initialize the problem}

  In addition to initializing the mass fractions, $X_k$, we also fill
  the auxiliary composition data, $Y_e$, $\bar{A}$, and $(B/A)$, using
  these $X_k$.

\item {\em Hydrodynamics update}

  The hydrodynamics update proceeds as normal, but with an advection
  equation for each of the auxiliary composition variables:
  \begin{align}
    \ddt{(\rho Y_e)} + \ddx{(\rho Y_e u)} &= 0 \\
    \ddt{(\rho \bar{A})} + \ddx{(\rho \bar{A} u)} &= 0 \\
    \ddt{[\rho (B/A)]} + \ddx{[\rho (B/A) u]} &= 0 
  \end{align}

\item {\em Reactive update}

  \begin{itemize}
  \item For a zone that is in NSE:

    \begin{itemize}
    \item use $\rho$, $T$, and $Y_e$ to call the table. This
      returns: $dY_e/dt$, $(B/A)_{\rm out}$, $\bar{A}_{\rm out}$,
      and the $(X_k)_{\rm out}$ from the NSE network grouped into the 19 isotopes
      we carry as part of aprox19.

    \item update $Y_e$:
    \begin{equation}
      (Y_e)_{\rm out} = (Y_e)_{\rm in} + \Delta t \frac{dY_e}{dt}
    \end{equation}

  \item update the auxiliary state value of $\bar{A}$ to be simply
    $\bar{A}_{\rm out}$ from the table and the mass fractions
    $X_k$ to be the $(X_k)_{\rm out}$ returned from the table.

@@

  \item For zones not in NSE:

@@

We use the following conditions to determine if we are in NSE:
\begin{align}
\rho &> \rho_\mathrm{nse} \\
T &> T_\mathrm{nse} \\
X(\isotm{C}{12}) &< A_\mathrm{nse} \\
X(\isotm{He}{4}) + \sum_{k \in {\rm Fe-group}} X_k &> B_\mathrm{nse} 
\end{align}
where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52}, \isot{Fe}{54}, and \isot{Ni}{56}.
